% =============================================================================
%  Bibliography
% =============================================================================

\begin{thebibliography}{36}

\bibitem{Shekhar2024}
R.~Shekhar, J.~Hunziker and Y.\,H.~Lim,
\textit{Integral equation method for microseismic wavefield modelling in
anisotropic elastic media},
Geophysical Prospecting, \textbf{72}, 403--423, 2024.

\bibitem{Korneev1993}
V.\,A.~Korneev and L.\,R.~Johnson,
\textit{Scattering of elastic waves by a spherical inclusion --- I.\ Theory
and numerical results},
Geophysical Journal International, \textbf{115}, 230--250, 1993.

\bibitem{Gritto1995}
R.~Gritto, V.\,A.~Korneev and L.\,R.~Johnson,
\textit{Low-frequency elastic-wave scattering by an inclusion: limits of
applications},
Geophysical Journal International, \textbf{120}, 677--692, 1995.

\bibitem{Pollitz1998}
F.\,F.~Pollitz,
\textit{Scattering of spherical elastic waves from a small-volume spherical
inclusion},
Geophysical Journal International, \textbf{134}, 390--408, 1998.

\bibitem{Pellegrini2011}
Y.-P.~Pellegrini, P.~Music and D.\,B.~Stout,
\textit{Mean-field T-matrix approach to elastic wave scattering by small and
point-like objects},
Wave Motion, \textbf{48}, 698--714, 2011. arXiv:1102.2822.

\bibitem{Waterman1976}
P.\,C.~Waterman,
\textit{Matrix theory of elastic wave scattering},
Journal of the Acoustical Society of America, \textbf{60}, 567--580, 1976.

\bibitem{Bostock1991}
M.\,G.~Bostock,
\textit{Surface wave scattering from 3-D obstacles},
Geophysical Journal International, \textbf{104}, 351--370, 1991.

\bibitem{BostockKennett1992}
M.\,G.~Bostock and B.\,L.\,N.~Kennett,
\textit{Multiple scattering of surface waves from discrete obstacles},
Geophysical Journal International, \textbf{108}, 52--70, 1992.

\bibitem{Jakobsen2012}
M.~Jakobsen,
\textit{T-matrix approach to seismic forward modelling in the acoustic
approximation},
Studia Geophysica et Geodaetica, \textbf{56}, 1--27, 2012.

\bibitem{JakobsenUrsin2015}
M.~Jakobsen and B.~Ursin,
\textit{Full waveform inversion in the frequency domain using direct iterative
T-matrix methods},
Journal of Geophysics and Engineering, \textbf{12}, 400--418, 2015.

\bibitem{JakobsenWu2016}
M.~Jakobsen and R.-S.~Wu,
\textit{Renormalized scattering series for frequency-domain waveform modelling
for strong velocity contrasts},
Geophysical Journal International, \textbf{206}, 880--899, 2016.

\bibitem{Hales2005}
T.\,C.~Hales,
\textit{A proof of the Kepler conjecture},
Annals of Mathematics, \textbf{162}, 1065--1185, 2005.

\bibitem{HashinShtrikman1962a}
Z.~Hashin and S.~Shtrikman,
\textit{On some variational principles in anisotropic and nonhomogeneous
elasticity},
Journal of the Mechanics and Physics of Solids, \textbf{10}, 335--342, 1962.

\bibitem{HashinShtrikman1962b}
Z.~Hashin and S.~Shtrikman,
\textit{A variational approach to the theory of the elastic behaviour of
polycrystals},
Journal of the Mechanics and Physics of Solids, \textbf{10}, 343--352, 1962.

\bibitem{HashinShtrikman1963}
Z.~Hashin and S.~Shtrikman,
\textit{A variational approach to the theory of the elastic behaviour of
multi-phase materials},
Journal of the Mechanics and Physics of Solids, \textbf{11}, 127--140, 1963.

\bibitem{Hashin1983}
Z.~Hashin,
\textit{Analysis of composite materials --- a survey},
Journal of Applied Mechanics, \textbf{50}, 481--505, 1983.

\bibitem{Willis1977}
J.\,R.~Willis,
\textit{Bounds and self-consistent estimates for the overall properties of
anisotropic composites},
Journal of the Mechanics and Physics of Solids, \textbf{25}, 185--202, 1977.

\bibitem{Willis1981}
J.\,R.~Willis,
\textit{Variational and related methods for the overall properties of
composites},
Advances in Applied Mechanics, \textbf{21}, 1--78, 1981.

\bibitem{ZellerDederichs1973}
R.~Zeller and P.\,H.~Dederichs,
\textit{Elastic constants of polycrystals},
Physica Status Solidi~(b), \textbf{55}, 831--842, 1973.

\bibitem{Gubernatis1977}
J.\,E.~Gubernatis, E.~Domany, J.\,A.~Krumhansl and M.~Huberman,
\textit{The Born approximation in the theory of the scattering of elastic
waves by flaws},
Journal of Applied Physics, \textbf{48}, 2812--2819, 1977.

\bibitem{Norris1985}
A.\,N.~Norris,
\textit{A differential scheme for the effective moduli of composites},
Mechanics of Materials, \textbf{4}, 1--16, 1985.

\bibitem{Kennett1983}
B.\,L.\,N.~Kennett,
\textit{Seismic Wave Propagation in Stratified Media},
Cambridge University Press, Cambridge, 1983.

\bibitem{Kennett1979}
B.\,L.\,N.~Kennett,
\textit{The suppression of surface multiples on seismic records},
Geophysical Prospecting, \textbf{27}, 584--600, 1979.

\bibitem{Foldy1945}
L.\,L.~Foldy,
\textit{The multiple scattering of waves.\ I.\ General theory of isotropic
scattering by randomly distributed scatterers},
Physical Review, \textbf{67}, 107--119, 1945.

\bibitem{Lax1952}
M.~Lax,
\textit{Multiple scattering of waves.\ II.\ The effective field in dense
systems},
Physical Review, \textbf{85}, 621--629, 1952.

\bibitem{SaadSchultz1986}
Y.~Saad and M.\,H.~Schultz,
\textit{GMRES: A generalized minimal residual algorithm for solving
nonsymmetric linear systems},
SIAM Journal on Scientific and Statistical Computing, \textbf{7}, 856--869, 1986.

\bibitem{vanderVorst1992}
H.\,A.~van der Vorst,
\textit{Bi-CGSTAB: A fast and smoothly converging variant of Bi-CG for the
solution of nonsymmetric linear systems},
SIAM Journal on Scientific and Statistical Computing, \textbf{13}, 631--644, 1992.

\bibitem{Rokhlin1985}
V.~Rokhlin,
\textit{Rapid solution of integral equations of classical potential theory},
Journal of Computational Physics, \textbf{60}, 187--207, 1985.

\bibitem{GreengardRokhlin1987}
L.~Greengard and V.~Rokhlin,
\textit{A fast algorithm for particle simulations},
Journal of Computational Physics, \textbf{73}, 325--348, 1987.

\bibitem{Fujiwara2000}
H.~Fujiwara,
\textit{The fast multipole method for solving integral equations of
three-dimensional topography and basin problems},
Geophysical Journal International, \textbf{140}, 198--210, 2000.

\bibitem{Chaillat2008}
S.~Chaillat, M.~Bonnet and J.-F.~Semblat,
\textit{A multi-level fast multipole BEM for 3-D elastodynamics in the
frequency domain},
Computer Methods in Applied Mechanics and Engineering, \textbf{197},
4233--4249, 2008.

\bibitem{Hackbusch1999}
W.~Hackbusch,
\textit{A sparse matrix arithmetic based on $\mathcal{H}$-matrices.\ I.\
Introduction to $\mathcal{H}$-matrices},
Computing, \textbf{62}, 89--108, 1999.

\bibitem{Boerm2003}
S.~B\"orm, L.~Grasedyck and W.~Hackbusch,
\textit{Introduction to hierarchical matrices with applications},
Engineering Analysis with Boundary Elements, \textbf{27}, 405--422, 2003.

\bibitem{Ewald1921}
P.\,P.~Ewald,
\textit{Die Berechnung optischer und elektrostatischer Gitterpotentiale},
Annalen der Physik, \textbf{369}, 253--287, 1921.

\bibitem{DederichsZeller1972}
P.\,H.~Dederichs and R.~Zeller,
\textit{Elastische Konstanten von Vielkristallen},
Berichte der Kernforschungsanlage J\"ulich, Nr.\ 877-FF, 1972.

\bibitem{Nestor1996}
T.\,M.~Nestor,
\textit{Seismic Wave Propagation in Complex Media: Summing the Multiple
Scattering Series},
PhD thesis, Australian National University, 1996.

\end{thebibliography}
