% LuaLaTeX document — compile with: lualatex cube_galerkin27.tex
\documentclass[11pt,a4paper]{article}

\usepackage{fontspec}
\setmainfont{Latin Modern Roman}
\setmonofont{Latin Modern Mono}

\usepackage{amsmath,amssymb,amsthm,bm}
\usepackage{booktabs}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!60!black,citecolor=green!50!black,urlcolor=blue!70!black}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{mathtools}
\usepackage{graphicx}  % \resizebox for long closed-form expressions

\newcommand{\D}{\Delta}
\newcommand{\order}{\mathcal{O}}
\newcommand{\bG}{\bm{G}}
\newcommand{\bs}{\bm{s}}
\newcommand{\br}{\bm{r}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\be}{\bm{\varepsilon}}
\newcommand{\bu}{\bm{u}}
\newcommand{\bphi}{\bm{\varphi}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\PV}{\mathrm{p.v.}}
\newcommand{\indV}{\chi_V}
\newcommand{\Lspace}{L^1_{\mathrm{loc}}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{A Rigorous Galerkin Closure for the Elastodynamic\\
Volume Integral Equation on a Cube\\
\large 27-Component Path-B Formulation}
\author{Research Notes}
\date{April 2026}

\begin{document}
\maketitle
\tableofcontents
\vspace{1em}

%======================================================================
\section{Motivation: differentiating under the integral fails}
%======================================================================

{\sloppy
The classical moment-method closure for the cube T-matrix begins from the
Lippmann--Schwinger (LS) equation\par}
\begin{equation}
  u_i(\br) \;=\; u_i^{0}(\br)
   + \omega^2 \!\int_V G_{ij}(\br - \br')\,\delta\rho(\br')\, u_j(\br')\,dV'
   - \!\int_V \partial'_k G_{in}(\br - \br')\,\delta c_{nklj}(\br')\,
   \partial'_l u_j(\br')\,dV',
   \label{eq:LS}
\end{equation}
where $G_{ij}$ is the elastostatic Green's tensor and $V \subset \R^3$ is the
inclusion. To obtain a closed finite-dimensional system one then formally
takes derivatives at the centre $\br = \mathbf{0}$:
\begin{align}
   u_i(\mathbf{0}) &= u_i^{0}(\mathbf{0}) + \omega^2\!\int_V G_{ij} \,\delta\rho\, u_j\, dV'
                       - \!\int_V \partial'_k G_{in}\,\delta c_{nklj}\,\partial'_l u_j\, dV', \nonumber\\
   \partial_p u_i(\mathbf{0}) &= \partial_p u_i^{0}(\mathbf{0})
                       - \omega^2\!\int_V \partial'_p G_{ij} \,\delta\rho\, u_j\, dV'
                       + \!\int_V \partial'_p \partial'_k G_{in}\,\delta c_{nklj}\,\partial'_l u_j\, dV', \label{eq:LS-derivs}\\
   \partial_q\partial_p u_i(\mathbf{0}) &= \partial_q\partial_p u_i^{0}(\mathbf{0})
                       + \omega^2\!\int_V \partial'_q\partial'_p G_{ij} \,\delta\rho\, u_j\, dV'
                       - \!\int_V \partial'_q\partial'_p\partial'_k G_{in}\,\delta c_{nklj}\,\partial'_l u_j\, dV',
   \nonumber
\end{align}
where the $r$--derivatives have been converted to $r'$--derivatives by translation
invariance $\partial_r G(\br - \br') = -\partial_{r'} G(\br - \br')$.

\medskip

The Leibniz rule used in passing from \eqref{eq:LS} to \eqref{eq:LS-derivs}
requires (DCT version) a dominating function $|\partial_r f(\br,\br')| \le g(\br')$
with $g \in L^{1}(V)$.  The static elastic kernel has the singularity
\begin{equation}
   G^{\text{stat}}_{ij}(\bs) \sim \frac{1}{|\bs|},
   \quad \partial' G \sim \frac{1}{|\bs|^2},
   \quad \partial'\partial' G \sim \frac{1}{|\bs|^3},
   \quad \partial'\partial'\partial' G \sim \frac{1}{|\bs|^4},
   \label{eq:singularities}
\end{equation}
so already at the level of $\partial' G$ \emph{no integrable dominating function exists} when $\br$ is in the support of the inclusion. Each of the equations in
\eqref{eq:LS-derivs} therefore makes only \emph{formal} sense; converting them into rigorous statements requires distributional regularization (Cauchy principal value, Hadamard finite part, and Eshelby-type delta corrections).

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
LS derivative order & worst kernel & status \\
\midrule
$0$ (body force)        & $1/|\bs|$   & $\Lspace$ \checkmark \\
$0$ (elastic kernel)    & $1/|\bs|^2$ & conditionally convergent (PV) \\
$1$ (body force)        & $1/|\bs|^2$ & PV \\
$1$ (elastic kernel)    & $1/|\bs|^3$ & \textbf{Eshelby delta required} \\
$2$ (body force)        & $1/|\bs|^3$ & \textbf{Eshelby delta required} \\
$2$ (elastic kernel)    & $1/|\bs|^4$ & \textbf{derivative-of-delta / Hadamard PF} \\
\bottomrule
\end{tabular}
\caption{Singular structure of the formal closure equations \eqref{eq:LS-derivs}.}
\label{tab:singularities}
\end{table}

%======================================================================
\section{Path B: Galerkin formulation on a polynomial trial space}
%======================================================================

We replace point evaluation by a Galerkin (weak) projection onto a
finite-dimensional polynomial trial space.  The resulting bilinear forms
involve only $G_{ij}$ itself (the singularity $1/|\bs|$ is locally
$L^{1}$ in $\R^{3}$), so \emph{every entry of the closure matrix is an
absolutely convergent integral}.  No principal values, no delta
corrections, no distributional bookkeeping.

\subsection{Trial space \texorpdfstring{$T_{27}$}{T27}}

Let $V = [-a,a]^{3}$.  Denote by $P_{2}(V)$ the space of real polynomials
in $\br = (r_1,r_2,r_3)$ of total degree $\le 2$, $\dim P_{2}(V) = 10$. The naive
vector trial space $(P_{2}(V))^{3}$ has dimension $30$.  We remove the
$3$ rigid-rotation modes by symmetrizing the linear part, giving the
$27$-dimensional subspace
\begin{equation}
   T_{27} \;=\; \Big\{\,\bphi(\br)\;\Big|\;
   \varphi_i(\br) = U_i + \varepsilon_{ai}\, r_a + \tfrac{1}{2}\,M_{abi}\, r_a r_b,\quad
   \varepsilon_{ai} = \varepsilon_{ia},\quad M_{abi} = M_{bai}\,\Big\}.
\end{equation}
The dimensions add as $3 + 6 + 18 = 27$, matching one displacement
monopole, the symmetric strain tensor, and the symmetric second-derivative
tensor of $\bu$ (equivalent to the strain gradient $\partial_k\varepsilon_{ij}$
through the kinematic identity
$\partial_k\partial_l u_m = \partial_k\varepsilon_{lm} + \partial_l\varepsilon_{km} - \partial_m\varepsilon_{kl}$).

\subsection{Galerkin LS}

\begin{definition}[Path-B closure]
The Path-B 27-component closure is the variational problem:
find $\bu^{h}\in T_{27}$ such that for all $\bphi \in T_{27}$,
\begin{equation}
   \langle\bphi,\bu^{h}\rangle_V
   \;=\;
   \langle\bphi,\bu^{0}\rangle_V
   + \omega^{2}\delta\rho\, \mathcal{B}_{\text{body}}[\bphi,\bu^{h}]
   - \delta c_{nklj}\, \mathcal{B}_{\text{el}}[\bphi,\bu^{h}],
   \label{eq:galerkin-LS}
\end{equation}
where the bilinear forms are
\begin{align}
\langle\bphi,\bu\rangle_V &:= \int_{V} \varphi_i(\br)\, u_i(\br)\,dV, \\[2pt]
\mathcal{B}_{\text{body}}[\bphi,\bu] &:= \int_{V}\!\!\int_{V} \varphi_i(\br)\,
            G_{in}(\br-\br')\, u_n(\br')\,dV\,dV', \\[2pt]
\mathcal{B}_{\text{el}}[\bphi,\bu] &:= \int_{V}\!\!\int_{V} \varphi_i(\br)\,
            \partial'_k G_{in}(\br-\br')\, \partial'_l u_j(\br')\,dV\,dV'.
\end{align}
\end{definition}

\subsection{Why this is well defined}

\begin{proposition}\label{prop:wellposed}
For polynomial $\bphi,\bu \in T_{27}$, every bilinear form in
\eqref{eq:galerkin-LS} is an absolutely convergent integral.  No
distributional regularization is needed.
\end{proposition}
\begin{proof}[Proof sketch]
The mass form is trivial because the integrand is a polynomial.
For $\mathcal{B}_{\text{el}}$, we use integration by parts on $\bu$
(legitimate because $\bu$ is a polynomial and the boundary term is a
2D integral against a smooth $L^1_{\text{loc}}$ kernel):
\begin{equation}
\mathcal{B}_{\text{el}}[\bphi,\bu]
=
-\int_{V}\!\!\int_V\!\varphi_i\,G_{in}\,\partial'_k\partial'_l u_j\,dV'\,dV
+ \int_{V}\!\!\oint_{\partial V}\!\varphi_i\,G_{in}\,\partial'_l u_j\,n_k\,dS'\,dV.
\end{equation}
{\sloppy
Both integrands now contain only $G_{in}(\br-\br') \sim 1/|\br-\br'|$ multiplied
by polynomials/smooth functions.  The 6-fold integral
$\int\!\!\int_{V\times V}1/|\br-\br'|\,dV\,dV'$ converges by the
Hardy--Littlewood--Sobolev inequality (or directly: in
$\R^{6}$ the singular set $\{\br = \br'\}$ has codimension $3$ and
$1/|\br-\br'|$ has order-1 singularity, hence is locally
$L^{1}$ in $\R^{6}$).  The 5-fold boundary integral is similar.
The body form is even simpler since no integration by parts is needed.\par}
\qed
\end{proof}

\subsection{Equivalence to point-collocation in the limit}

The classical Path-A closure (point evaluation at $\br = \mathbf{0}$) is
the formal limit of a Petrov--Galerkin scheme with Dirac-delta test
functions.  Both Path A and Path B converge to the exact volume integral
equation as the polynomial degree $\to \infty$; at finite truncation
their stiffness matrices differ but encode the same physical operator.
Path B is preferable for analysis because it is symmetric, energy
conserving, and free of singular regularizations.

%======================================================================
\section{Center-of-mass / relative coordinates}
\label{sec:cm-relative}
%======================================================================

The 6-fold integrals in \eqref{eq:galerkin-LS} are reduced to 3-fold
integrals against the \emph{cube autocorrelation} via the \emph{centre-of-mass and relative coordinate} change of variables (also called \emph{sum-and-difference variables} in the Eshelby/Mura literature, or \emph{Jacobi coordinates} in the simplest 2-body case):
\begin{equation}
   \bxi = \tfrac{1}{2}(\br + \br'), \qquad \bs = \br - \br',
   \qquad \br = \bxi + \tfrac{\bs}{2}, \quad \br' = \bxi - \tfrac{\bs}{2},
   \qquad d\br\,d\br' = d\bxi\,d\bs.
\end{equation}
The constraint $\br,\br' \in V$ becomes $\bxi \in V \cap (V + \bs/2)\cap (V - \bs/2)$, which for the cube $V = [-a,a]^{3}$ simplifies to a centred box of half-lengths
$b_i(\bs) = a - |s_i|/2$:
\begin{equation}
   \bxi \in B(\bs) := \prod_{i=1}^{3} \big[-(a - |s_i|/2),\, a - |s_i|/2\big],
   \qquad \text{valid when } |s_i| < 2a.
\end{equation}
The 6-fold integral becomes
\begin{equation}
   \int_{V}\!\!\int_{V} P(\br)\,K(\br - \br')\,Q(\br')\,d\br\,d\br'
   \;=\; \int_{[-2a,2a]^{3}}\! K(\bs)
        \underbrace{\Big[\int_{B(\bs)}\!P\!\big(\bxi+\tfrac{\bs}{2}\big)\,Q\!\big(\bxi-\tfrac{\bs}{2}\big)\,d\bxi\Big]}_{\text{polynomial in } (\bs, a)}\,d\bs.
   \label{eq:cm-relative}
\end{equation}
For polynomial $P,Q$ the inner integral is elementary: every $\xi_i^{p}$
integrates over $[-b_i, b_i]$ to $\frac{2 b_i^{p+1}}{p+1}$ when $p$ is even and to $0$
when $p$ is odd, giving a polynomial in
$(b_1,b_2,b_3) = (a-|s_1|/2,\,a-|s_2|/2,\,a-|s_3|/2)$.

\medskip

\noindent\textbf{Cube autocorrelation.} For $P = Q = 1$ the inner integral
is exactly the autocorrelation of the cube indicator,
\begin{equation}
   (\indV \star \indV)(\bs) \;=\; \prod_{i=1}^{3} \big(2a - |s_i|\big)_{+},
\end{equation}
the 3-D \emph{tent function} $\Lambda_V(\bs)$.

%======================================================================
\section{Master integrals}
%======================================================================

After the centre-of-mass/relative reduction, every entry of the
$27\times 27$ Galerkin matrix is a finite linear combination of the
following \emph{master integrals} over the doubled cube
$\Omega = [-2a,2a]^{3}$:
\begin{align}
   \mathcal{I}_{p,q,r}(a) &:= \int_{\Omega}\!\frac{s_1^{p}\,s_2^{q}\,s_3^{r}}{|\bs|}\,d\bs, \label{eq:Imaster}\\
   \mathcal{J}^{(ij)}_{p,q,r}(a) &:= \int_{\Omega}\!\frac{s_i s_j\,s_1^{p}\,s_2^{q}\,s_3^{r}}{|\bs|^{3}}\,d\bs.
   \label{eq:Jmaster}
\end{align}
For the $9$-component (current) closure only $p+q+r \le 3$ is needed; the
$27$-component closure requires $p+q+r \le 5$.

By the parity $\bs \to -\bs$ on each axis separately, $\mathcal{I}_{p,q,r}$
vanishes whenever any of $p,q,r$ is odd, and $\mathcal{J}^{(ij)}_{p,q,r}$
vanishes unless the multiset $\{i,j,p\text{-copies of }1,q\text{-copies of }2,r\text{-copies of }3\}$ has even count for each axis.  This collapses
the formally $\binom{8}{3} \cdot 3 = 168$ entries down to $\sim 20$ inequivalent integrals after cubic symmetry.

\subsection{Tier-1 results (computed analytically)}\label{sec:tier1}

The basic integral $\mathcal{I}_{0,0,0}(a)$ is the dimensionless cube
constant $g_0$ scaled by $a^{2}$.  Mathematica's \texttt{Integrate}
returns it in the form
\begin{equation}
   \int_{[-1,1]^{3}}\!\frac{1}{|\bu|}\,du \;=\; -2\Big(\pi + \log\!\big(1351 - 780\sqrt{3}\big)\Big),
\end{equation}
which equals the Pell-stable form used in the existing
\texttt{CubeAnalytic.wl} package via the identity
$1351 - 780\sqrt{3} = (70226 + 40545\sqrt{3})^{-2/3}$ (both sides are Pell
units of $\mathbb{Z}[\sqrt{3}]$, satisfying $1351^{2} - 3\cdot 780^{2} = 1$
and $70226^{2} - 3\cdot 40545^{2} = 1$). Numerical agreement to $\sim\!78$ digits has been verified.

\input{cube_galerkin27_results.tex}

%======================================================================
\section{Validation plan: Path-A $\Leftrightarrow$ Path-B for the 9-component subsystem}
%======================================================================

The existing \texttt{CubeAnalytic.wl} package implements Path A with
analytical Eshelby-delta corrections (the $(5\sqrt{3} - 2\pi)$ block from
the old \texttt{GreensTensorScatteringPaper\_Old.nb}).  As a correctness
test for the framework presented here, we will:
\begin{enumerate}[label=(V\arabic*)]
\item Compute the Path-B mass and stiffness matrices on the 9-dimensional
      subspace $T_9 \subset T_{27}$ spanned by constants and symmetric
      linear polynomials.
\item Solve the resulting $9\times 9$ generalized eigenproblem for the
      effective contrasts $\Delta\rho^{*}, \Delta\lambda^{*}, \Delta\mu^{*}$.
\item {\sloppy Compare to the Path-A values produced by
      \texttt{compute\_\allowbreak cube\_\allowbreak tmatrix} in
      \texttt{cubic\_\allowbreak scattering/\allowbreak effective\_\allowbreak contrasts.py}.\par}
\end{enumerate}
Path A and Path B do not need to give identical matrix entries (they
correspond to different finite-dimensional projections of the same
operator), but the effective contrasts must agree to leading order in
the polynomial truncation.  Any discrepancy quantifies the truncation
error and bounds the systematic error of Path A.

%======================================================================
\section{New integrals required for the 27-component closure}
%======================================================================

Beyond the constants used in the 9-component closure
($\mathcal{I}_{0,0,0}$, $\mathcal{J}^{(11)}_{0,0,0}$,
$\mathcal{J}^{(11)}_{2,0,0}$, etc.), the $27$-component closure
introduces master integrals with total monomial degree up to $5$.
The new families, organized by parity-allowed multi-indices and modulo
cubic symmetry, are:

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
$(p,q,r)$ class & rank & needed for \\
\midrule
$(2,2,0),\;(4,0,0)$         & 4 & $\langle \bphi^{(\text{lin})},G\bphi^{(\text{lin})}\rangle$ block \\
$(2,2,2),\;(4,2,0),\;(6,0,0)$ & 6 & $\langle \bphi^{(\text{quad})},G\bphi^{(\text{lin})}\rangle$ block \\
$(4,2,2),\;(4,4,0),\;(6,2,0),\;(8,0,0)$ & 8 & $\langle \bphi^{(\text{quad})},G\bphi^{(\text{quad})}\rangle$ block \\
\bottomrule
\end{tabular}
\caption{New cubic-symmetric master integral families needed for the
27-component Path-B closure. Each row corresponds to one block of the
$27\times 27$ stiffness matrix.}
\label{tab:new-masters}
\end{table}

These integrals will be derived analytically in Section~\ref{sec:tier1}
and the subsequent stages of \texttt{CubeGalerkin27.wl}.

\input{cube_tmatrix_closedform.tex}

\bibliographystyle{plain}
\begin{thebibliography}{9}
\bibitem{mura1987} T.\ Mura, \emph{Micromechanics of Defects in Solids},
   2nd ed., Martinus Nijhoff (1987).
\bibitem{khachaturyan} A.\ G.\ Khachaturyan, \emph{Theory of Structural
   Transformations in Solids}, Wiley (1983).
\bibitem{costabel-stephan} M.\ Costabel and E.\ Stephan, ``Boundary
   integral equations for mixed boundary value problems in polygonal
   domains and Galerkin approximation'', \emph{Banach Center Publ.}
   \textbf{15} (1985), 175--251.
\bibitem{hsiao-wendland} G.\ C.\ Hsiao and W.\ L.\ Wendland,
   \emph{Boundary Integral Equations}, Springer (2008).
\end{thebibliography}

\end{document}
